---
title: "WG Ch7 & Ch8 Exercise Solutions"
author: Rachel Lee
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
library(kableExtra)
options(knitr.table.format = "latex")
```

Last compiled: `r Sys.time()`

<br>

*For Week 5: read WG chapters 6-8.*  


<br>
<br>

# Chapter 7 
<br>

### 7.3 Exercises

#### 1. Explore the distribution of each of the `x`, `y`, and `z` variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.

```{r}
library(tidyverse)
summary(select(diamonds, x, y, z))
```

The summary of the selected variables provides a good initial scan of our dataset. Note that `x` variable is the diamond length in mm, `y` is width in mm, and `z` is depth in mm. The typical values of `x` and `y` are larger than `z`, with `x` and `y` having inter-quartile ranges of 4.7–6.5, while `z` has an inter-quartile range of 2.9–4.0.

```{r}
diamonds %>% arrange(desc(x)) %>% top_n(5)
diamonds %>% arrange(desc(y)) %>% top_n(5)
diamonds %>% arrange(desc(z)) %>% top_n(5)
```

The top 5 diamonds with largest `x`, `y`, and `z` are shown above. We suspect that `y = 58.90` may be a typo. We can attempt to see the outliers visually:

```{r}
ggplot(diamonds, aes(x = x, y = y)) +
  geom_point(alpha = 0.5)
ggplot(diamonds, aes(x = x, y = z)) +
  geom_point(alpha = 0.5)
ggplot(diamonds, aes(x = y, y = z)) +
  geom_point(alpha = 0.5)
```
<br>
<br>

#### 2. Explore the distribution of `price`. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)









