---
title: "WG Ch4 Exercise Solutions"
author: Rachel Lee
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
library(kableExtra)
options(knitr.table.format = "latex")
```

Last compiled: `r Sys.time()`

<br>

*For Week 4: read WG chapters 4-5.*  
*Thank you Jean for catching some typos last week.*

<br>
<br>

# Chapter 4
<br>

### 4.4 Exercises

#### 1. Why does this code not work?
```{r}
my_variable <- 10
my_var覺able
#> Error in eval(expr, envir, enclos): object 'my_var覺able' not found
```

If you look carefully, `my_variable` is different from `my_var覺able`. Reading the error messages help: `object 'my_var覺able' not found` is very clear on where R is finding the error.  

<br>
<br>

#### 2. Tweak each of the following R commands so that they run correctly.

The corrected code should look like this:

```{r}
library(tidyverse)

# spelling mistake; change to data
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# set cyl == 8
filter(mpg, cyl == 8)

# data name is diamonds, not diamond
filter(diamonds, carat > 3)
```

<br>
<br>

#### 3. Press Alt + Shift + K. What happens? How can you get to the same place using the menus?  

This gives a menu with keyboard shortcuts. You can also access htem in the menu under `Tools -> Keyboard Shortcuts Help`.




