---
title: "WG Ch7 & Ch8 Exercise Solutions"
author: Rachel Lee
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
library(kableExtra)
options(knitr.table.format = "latex")
```

Last compiled: `r Sys.time()`

<br>

*For Week 5: read WG chapters 6-8.*  


<br>
<br>

# Chapter 7 
<br>

### 7.3 Exercises

#### 1. Explore the distribution of each of the `x`, `y`, and `z` variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.

```{r}
library(tidyverse)
summary(select(diamonds, x, y, z))
```

The summary of the selected variables provides a good initial scan of our dataset. Note that `x` variable is the diamond length in mm, `y` is width in mm, and `z` is depth in mm. The typical values of `x` and `y` are larger than `z`, with `x` and `y` having inter-quartile ranges of 4.7–6.5, while `z` has an inter-quartile range of 2.9–4.0.

```{r}
diamonds %>% arrange(desc(x)) %>% top_n(5)
diamonds %>% arrange(desc(y)) %>% top_n(5)
diamonds %>% arrange(desc(z)) %>% top_n(5)
```

The top 5 diamonds with largest `x`, `y`, and `z` are shown above. We suspect that `y = 58.90` may be a typo. We can attempt to see the outliers visually:

```{r}
ggplot(diamonds, aes(x = x, y = y)) +
  geom_point(alpha = 0.5)
ggplot(diamonds, aes(x = x, y = z)) +
  geom_point(alpha = 0.5)
ggplot(diamonds, aes(x = y, y = z)) +
  geom_point(alpha = 0.5)
```
<br>
<br>

#### 2. Explore the distribution of `price`. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)

```{r}
ggplot(filter(diamonds, price < 10000), aes(x = price)) +
  geom_histogram(binwidth = 10, center = 0)
```

From our graph above, We see something is off. To get a close-up view, we restrict the `price` range:
```{r}
ggplot(filter(diamonds, price < 2000), aes(x =price)) +
  geom_histogram(binwidth = 10, center = 0)
```

There are no diamonds with a price between $1,455 and $1,545.  

<br>

Are there anything else you notice?  

<br>
<br>

#### 3. How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?

```{r}
diamonds %>%
  filter(carat >= 0.99, carat <= 1) %>%
  count(carat)
```

There are 23 diamonds that are 0.99 carat weight, while 1558 diamonds with 1.00 carat weight. Maybe diamond sellers round up the carat? I am not too sure.  

<br>
<br>

#### 4. Compare and contrast `coord_cartesian()` vs `xlim()` or `ylim()` when zooming in on a histogram. What happens if you leave `binwidth` unset? What happens if you try and zoom so only half a bar shows?

```{r}
ggplot(diamonds) +
  geom_histogram(mapping = aes(x = price)) +
  coord_cartesian(xlim = c(100, 5000), ylim = c(0, 3000))

ggplot(diamonds) +
  geom_histogram(mapping = aes(x = price)) +
  xlim(100, 5000) +
  ylim(0, 3000)
```

The `coord_cartesian()` function zooms in on the area specified by the limits, after having calculated and drawn the geoms. Since the histogram bins have already been calculated, it is unaffected.  

However, the `xlim()` and `ylim()` functions influence actions before the calculation of the stats related to the histogram. Thus, any values outside the x- and y-limits are dropped before calculating bin widths and counts. This can influence how the histogram looks.

<br>
<br>


### 7.4 Exercises

#### 1. What happens to missing values in a histogram? What happens to missing values in a bar chart? Why is there a difference?  

Missing values are removed when the number of observations in each bin are calculated and warning sign is generated.  

<br>
<br>

#### 2. What does `na.rm = TRUE` do in `mean()` and `sum()`?

This option removes `NA` values from the vector prior to calculating the mean and sum.  

<br>
<br>





