---
title: "WG Ch4 & Ch5 Exercise Solutions"
author: Rachel Lee
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
library(kableExtra)
options(knitr.table.format = "latex")
```

Last compiled: `r Sys.time()`

<br>

*For Week 4: read WG chapters 4-5.*  

*Thank you Jean for catching some typos last week.*

<br>
<br>

# Chapter 4
<br>

### 4.4 Exercises

#### 1. Why does this code not work?
```{r}
my_variable <- 10
my_varıable
#> Error in eval(expr, envir, enclos): object 'my_varıable' not found
```

If you look carefully, `my_variable` is different from `my_varıable`. Reading the error messages help: `object 'my_varıable' not found` is very clear on where R is finding the error.  

<br>
<br>

#### 2. Tweak each of the following R commands so that they run correctly.

The corrected code should look like this:

```{r}
library(tidyverse)

# spelling mistake; change to data
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# set cyl == 8
filter(mpg, cyl == 8)

# data name is diamonds, not diamond
filter(diamonds, carat > 3)
```

<br>
<br>

#### 3. Press Alt + Shift + K. What happens? How can you get to the same place using the menus?  

This gives a menu with keyboard shortcuts. You can also access htem in the menu under `Tools -> Keyboard Shortcuts Help`.

<br>
<br>

# Chapter 4
<br>

# Chapter 5

### 5.2 Exercises
```{r}
library(nycflights13)
library(tidyverse)
```
#### 1. Find all flights that
1. Had an arrival delay of two or more hours
```{r}
filter(flights, arr_delay >= 120)
```

2. Flew to Houston (IAH or HOU)]
```{r}
filter(flights, dest == "IAH" | dest == "HOU")
```
If you have been practicing on DataCamp, you may have this code below that also produces the same output:
```{r}
filter(flights, dest %in% c("IAH", "HOU"))
```

3. Were operated by United, American, or Delta
```{r}

```

4. Departed in summer (July, August, and September)
5. Arrived more than two hours late, but didn’t leave late
6. Were delayed by at least an hour, but made up over 30 minutes in flight
7. Departed between midnight and 6am (inclusive)

