---
title: "Exercise Solutions"
author: Rachel Lee
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

Last compiled: `r Sys.Date()`

# Chapter 3


### 3.2.4 Exercises

#### 1. Run `ggplot(data = mpg)`. What do you see?
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)

ggplot(data = mpg)
```

This is an empty plot since no layers were added with geom function.


#### 2. How many rows are in `mtcars`? How many columns?
```{r}
nrow(mtcars)
ncol(mtcars)
```

There are `r nrow(mtcars)` rows and `r ncol(mtcars)` columns in `mtcars`.


#### 3. What does the `drv` variable describe? Read the help for `?mpg` to
find out.
```{r}
# ?mpg
```

`drv` is the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd.


#### 4. Make a scatterplot of `hwy` versus `cyl`.
```{r}
ggplot(mpg, aes(x = cyl, y = hwy)) +
  geom_point()
```


#### 5. What happens if you make a scatterplot of `class` versus `drv`?
Why is the plot not useful?
```{r}
ggplot(mpg, aes(x = class, y = drv)) +
  geom_point()
```

A scatter plot is not a useful here because both `drv` and `class` are categorical variables. Scatterplots are commonly used to visualize two continuous variables using visual marks mapped to a two-dimensional Cartesian space. Hence, the scatterplot here shows the limitation of managing two categorical variables. This is largely because categorical variables are typically stored as a small number of values. Within the `mpg` dataset, we can see that `class` variables holds 7 types of cars, and `cyl` holds 4 types of cylinders. Therefore, there will only be 21 points mapped on the scatterplot.

```{r}
mpg %>% distinct(class) # or unique(mpg$class)
mpg %>% distinct(cyl)   # or unique(mpg$cyl)
```


### 3.3.1 Exercises

#### 1. What’s gone wrong with this code? Why are the points not blue?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

The `color` argument is defined as part of the `mapping`. Therefore, it is being argued as an aesthetic. This can be corrected as below:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

Or,

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_point(colour = "blue")
```


#### 2. Which variables in `mpg` are categorical? Which variables are continuous? (Hint: type `?mpg` to read the documentation for the dataset). How can you see this information when you run `mpg`?

```{r}
head(mpg, 10) # or glipse(mpg)
```
My initial screening is that Variables with `<chr>` are categorical, those with `<dbl>` and `<int>` are continuous. But if you take a closer look at the dataset, `cyl`, `disp` are categorical. `year`, `city`, and `hwy`discrete variables, but we can decide to treat it as continuous variables.


#### 3. Map a continuous variable to `color`, `size`, and `shape`. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r}
ggplot(mpg, aes(x = cty, y = hwy, color = drv)) +
  geom_point() +
  labs(title = "mpg dataset: hwy vs. cty", 
       caption = "mpg dataset", 
       x = "Highway miles per gallon", 
       y = "city miles per gallon",
       color = "Type of drive train")
```

It's intuitive that `cty` and `hwy` will show a positive linear relationship. The `color` mapping is set to `drv`, which is a categorical variable.  


```{r}
ggplot(mpg, aes(x = cty, y = hwy, color = year)) +
  geom_point() +
  labs(title = "mpg dataset: hwy vs. cty", 
       caption = "mpg dataset", 
       x = "Highway miles per gallon", 
       y = "city miles per gallon",
       color = "Year")
```




# Chapter 4

## Second level header

You can add more headers by adding more hashtags. These won't be put into the table of contents

### third level header

Here's an even lower level header

# My second post (note the order)

2018 | 7 | 23 Last compiled: `r Sys.Date()`

I'm writing this tutorial going from the top down. And, this is how it will be printed. So, notice the second post is second in the list. If you want your most recent post to be at the top, then make a new post starting at the top. If you want the oldest first, do, then keep adding to the bottom

# Adding R stuff

So far this is just a blog where you can write in plain text and serve your writing to a webpage. One of the main purposes of this lab journal is to record your progress learning R. The reason I am asking you to use this process is because you can both make a website, and a lab journal, and learn R all in R-studio. This makes everything really convenient and in the sam place. 

So, let's say you are learning how to make a histogram in R. For example, maybe you want to sample 100 numbers from a normal distribution with mean = 0, and standard deviation =1, and then you want to plot a histogram. You can do this right here by using an r code block, like this:

```{r}
samples <- rnorm(100, mean=0, sd=1)
hist(samples)
```

When you knit this R Markdown document, you will see that the histogram is printed to the page, along with the R code. This document can be set up to hide the R code in the webpage, just delete the comment (hashtag), from the cold folding option in the yaml header up top. For purposes of letting yourself see the code, and me see the code, best to keep it the way that it is. You learn all of these things and more can be customized in each R code block.

# The big idea

Use this lab journal to record what you do in R. This way I will be able to see what you are doing and help you along the way. You will also be creating a repository of all the things you do. You can make posts about everything. Learning specific things in R (project unrelated), and doing things for the project that we will discuss at the beginning of the Fall semester. You can get started now by fiddling around with googling things, and trying stuff out in R. I've placed some helpful starting links in the links page on this website

# What can you do right now by yourself?

It's hard to learn programming when you don't have specific problems that you are trying to solve. Everything just seems abstract.

I wrote an [introductory programming book that introduces R](https://crumplab.github.io/programmingforpsych/), and gives some [concrete problems for you to solve](https://crumplab.github.io/programmingforpsych/programming-challenges-i-learning-the-fundamentals.html). 

To get the hang of journaling and solving the problems to learn programming, my suggestion is that you use this .Rmd file to solve the problems. It would look like this:

# Problem 1

Do simple math with numbers, addition, subtraction, multiplication, division

```{r}
1+2
2*5
5/3
(1+6+4)/5

```

# Problem 2

Put numbers into variables, do simple math on the variables

```{r}
a<-1
b<-2
a+b

d<-c(1,2,3)
e<-c(5,6,7)
d+e
d*e
d/e

```

# Problem 3

Write code that will place the numbers 1 to 100 separately into a variable using for loop. Then, again using the seq function.

```{r}
# for loop solution
# i becomes the number 1 to 100 at each step of the loop


a <- length(100) # make empty variable, set length to 100
for (i in 1:100){
  a[i] <-i #assigns the number in i, to the ith index of a
}

print(a)

# for loop solution #2

a<-c() #create empty variable using combine command
for (i in 1:100){
  a<-c(a,i) # keeps combining a with itself and the new number in i
}
print(a)

# seq solution
a <- seq(1,100,1) # look up help for seq using ?seq() in console
print(a)

```

# Replace this with problem 4

And keep going. Try to solve the problems with different scripts that provide the same solution. Good luck, Happy coding.
