---
title: "WG Ch4 & Ch5 Exercise Solutions"
author: Rachel Lee
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
library(kableExtra)
options(knitr.table.format = "latex")
```

Last compiled: `r Sys.time()`

<br>

*For Week 4: read WG chapters 4-5.*  

*Thank you Jean for catching some typos last week :)*

<br>
<br>

# Chapter 4 
<br>

### 4.4 Exercises

#### 1. Why does this code not work?
```{r}
my_variable <- 10
my_varıable
#> Error in eval(expr, envir, enclos): object 'my_varıable' not found
```

If you look carefully, `my_variable` is different from `my_varıable`. Reading the error messages help: `object 'my_varıable' not found` is very clear on where R is finding the error.  

<br>
<br>

#### 2. Tweak each of the following R commands so that they run correctly.

The corrected code should look like this: 

```{r}
library(tidyverse)

# spelling mistake; change to data
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# set cyl == 8
filter(mpg, cyl == 8)

# data name is diamonds, not diamond
filter(diamonds, carat > 3)
```

<br>
<br>

#### 3. Press Alt + Shift + K. What happens? How can you get to the same place using the menus?  

This gives a menu with keyboard shortcuts. You can also access htem in the menu under `Tools -> Keyboard Shortcuts Help`.

<br>
<br>
<br>

# Chapter 5

### 5.2 Exercises
```{r}
library(nycflights13)
library(tidyverse)
```
#### 1. Find all flights that

1. Had an arrival delay of two or more hours
```{r}
filter(flights, arr_delay >= 120)
```
<br>

2. Flew to Houston (IAH or HOU)
```{r}
filter(flights, dest == "IAH" | dest == "HOU")
```

If you have been practicing on DataCamp, you may have this code below that also produces the same output:
```{r}
filter(flights, dest %in% c("IAH", "HOU"))
```
<br>

3. Were operated by United, American, or Delta  

You first need to look up the abbreviations for airlines.
```{r}
airlines
```

Now using the carrier codes, we can find those operated by the three specified airlines:
```{r}
filter(flights, carrier %in% c("AA", "DL", "UA"))
```
<br>

4. Departed in summer (July, August, and September)

Any one of the four codes will produce the same output. *Note: my outputs are suppressed using the rmd function `results = FALSE`.*
```{r results = FALSE}
filter(flights, month >= 7 & month <= 9)
filter(flights, month %in% 7:9)
filter(flights, month >= 7, month <= 9)
filter(flights, month == 7 | month == 8 | month == 9)
```
<br>

5. Arrived more than two hours late, but didn’t leave late
```{r}
filter(flights, arr_delay > 120 & dep_delay <= 0)
```
<br>

6. Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)
```

<br>

7. Departed between midnight and 6am (inclusive)
```{r}
filter(flights, dep_time == 2400 | dep_time <= 600)
```
<br>
<br>

#### 2. Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?

`?between()` shows the usage. It takes 3 arguments: `x`, `left`, and `right`. Try running `between(1:12, 7, 9)`:
```{r}
between(1:12, 7, 9)
```

We can use this function to simploy the previous question:
```{r}
filter(flights, between(month, 7, 9))
```

<br>
<br>

#### 3. How many flights have a missing `dep_time`? What other variables are missing? What might these rows represent?

We can use `is.na()` function. DataCamp covers this function with numerous examples.
```{r}
flights %>% filter(is.na(dep_time))
flights %>% filter(is.na(dep_time)) %>% count()
```
From the given output table, `arr_time` and `dep_delay` are also missing. There are 8255 flights with missing `dep_time`.  

<br>
<br>

#### 4. Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing? Why is `FALSE & NA` not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)

```{r}
NA^0
NA | TRUE
FALSE & NA
NA * 0
```
`NA^0` results in 1 for the same reason that any number raised to 0th power is 1.  
`NA | TRUE` because `|` means "OR". So, anything or `TRUE` is `TRUE`.  
`FALSE & NA` is `FALSE` because `&` means "AND". Anything AND `FALSE` is `FALSE`.  
`NA * 0` is `NA` because `NA` contains undefined values, such as negative or positive infinity.  

<br>
<br>

### 5.3 Exercises

#### 1. How could you use `arrange()` to sort all missing values to the start? (Hint: use `is.na()`).

We will use the `flights` dataset.

```{r}
flights %>% arrange(is.na(dep_time))
```
It look slike `arrange(is.na(dep_time))` is placing those with `NA` at the very last. This is because R always handles `TRUE > FALSE`. Therefore, if we use the `desc()` function around `is.na()`, it should give us the correct result.

```{r}
flights %>% arrange(desc(is.na(dep_time)))
```

<br>
<br>

#### 2. Sort flights to find the most delayed flights. Find the flights that left earliest.

```{r}
flights %>% arrange(desc(dep_delay))
```

Most delayed flight was HA51 from JFK to HNL. It was delayed for 1301 minutes.  

Finding the flights that left the earliest is as follows:

```{r}
flights %>% arrange(dep_delay)
```

<br>
<br>

#### 3. Sort flights to find the fastest (highest speed) flights.

```{r}
flights %>% arrange(desc(distance/air_time))
```

We can find the speed of flights by dividing `distance` by `air_time`.

<br>
<br>

#### 4. Which flights travelled the farthest? Which travelled the shortest?

Finding the flights that traveled the farthest:

```{r}
flights %>% arrange(desc(distance))
```

Flights that traveled the shortest:

```{r}
flights %>% arrange(distance)
```
<br>
<br>

### 5.4 Exercises

#### 1. Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from `flights`.

```{r}
flights %>% select(dep_time, dep_delay, arr_time, arr_delay)
flights %>% select(4, 6, 7, 9) #specifying column numbers
flights %>% select(starts_with("dep_"), starts_with("arr_"))
```

If you can think of anymore, please email me!

<br>
<br>

#### 2. What happens if you include the name of a variable multiple times in a `select()` call?

We can try this.
```{r}
flights %>% select(year, year, day, day, day)
```

So it looks like R ignores the duplicated commands and selects the called column once.

<br>
<br>

#### 3. What does the `any_of()` function do? Why might it be helpful in conjunction with this vector?

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
```

`?any_of()` documentation reads that the function will select variables with a character vector rather than unquoted variable name arguments. Below is how it can be used using the `vars` called earlier.

```{r}
flights %>% select(one_of(vars))
```

<br>
<br>

#### 4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
select(flights, contains("TIME"))
```

This might be surprising because the default behavior for contains() is to ignore upper or lower cases. You can find this by reading the documetation of `contains()`. We can manually turn this off by adding `ignore.case = FALSE`:

```{r}
select(flights, contains("TIME", ignore.case = FALSE))
```

<br>
<br>

### 5.5 Exercises















